---
title: "Lab_9a_EvoGeno_Ben_Earle"
author: "Ben Earle"
date: "10/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
```

```{r, include=FALSE}
library(stringr)
```

# 14.3 Matching patterns with regular expressions

## 14.3.1 Basic matches
### The simplest patterns match exact strings:
```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
```

### The next step up in complexity is ., which matches any character (except a newline):
```{r}
str_view(x, ".a.")
```

### But if “.” matches any character, how do you match the character “.”? 

```{r}
x <- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```
### If \is used as an escape character in regular expressions, how do you match a literal \?
```{r}
x <- "a\\b"
writeLines(x)
```

```{r}
str_view(x, "\\\\")
```

## 14.3.1.1 Exercises
## 1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".
### None of these strings match a \ because you need 4 backslashes to match one backslash.m 
## 2. How would you match the sequence "'\?
```{r}
x <- "\\"
```
```{r}
str_view(x, "\\\\")
```
## 3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string?
```{r}
x <- "\\..\\..\\.."
```
```{r}
str_view(x, ".\\\\..\\\\..\\\\...")
```

## 14.3.2 Anchors

### By default, regular expressions will match any part of a string. It’s often useful to anchor the regular expression so that it matches from the start or end of the string. You can use:

### ^ to match the start of the string.
### $ to match the end of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
```
```{r}
str_view(x, "a$")
```

### To force a regular expression to only match a complete string, anchor it with both ^ and $:
```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "apple")
```
```{r}
str_view(x, "^apple$")
```

## 14.3.2.1 Exercises
### 1. How would you match the literal string "$^$"?
```{r}
x <- "$^$"
```
```{r}
str_view(x, ".$^$.")
```
### 2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:
### Start with “y”.

```{r}
str_view(stringr::words, "^y", match = TRUE)
```
### End with “x”
```{r}
str_view(stringr::words, "x$", match = TRUE)
```
### Are exactly three letters long. (Don’t cheat by using str_length()!)
```{r}
str_view(stringr::words, "^...$", match = TRUE)
```

### Have seven letters or more.
```{r}
str_view(stringr::words, ".......", match = TRUE)
```

## 14.3.3 Character classes and alternatives
### There are a number of special patterns that match more than one character. You’ve already seen ., which matches any character apart from a newline. There are four other useful tools:
### \d : matches any digit.
### \s : matches any whitespace (e.g. space, tab, newline).
### [abc] : matches a, b, or c.
### [^abc] : matches anything except a, b, or c.

### Look for a literal character that normally has special meaning in a regex
```{r}
x <- c("abc", "a.c", "a*c", "a c")
str_view(x, "a[.]c")
```

```{r}
str_view(x, ".[*]c")
```
```{r}
str_view(x, "a[ ]")
```

### This works for most (but not all) regex metacharacters: $ . | ? * + ( ) [ {. Unfortunately, a few characters have special meaning even inside a character class and must be handled with backslash escapes: ] \ ^ and -.

```{r}
x <- c("grey", "gray")
str_view(x, "gr(e|a)y")
```

## 14.3.3.1 Exercises
## 1 Create regular expressions to find all words that:
### Start with a vowel.
```{r}
str_view(stringr::words, "^[aeiou]", match = TRUE)
```

### That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r}
str_view(stringr::words, "[aeiou]", match = FALSE)
```

### End with ed, but not with eed.
```{r}
str_view(stringr::words, "[^e]ed$", match = TRUE)
```
### End with ing or ise.
```{r}
str_view(stringr::words, "i(ng|se)$", match = TRUE)
```
## 2 Empirically verify the rule “i before e except after c”.
```{r}
str_view(stringr::words, "(cei|[^c]ie)", match = TRUE)
```
```{r}
str_view(stringr::words, "(cie|[^c]ei)", match = TRUE)
```
## 3 Is “q” always followed by a “u”?
```{r}
str_view(stringr::words, "q[^u]", match = TRUE)
```
### in the stringr::words dataset, yes.

## 4 Write a regular expression that matches a word if it’s probably written in British English, not American English.
```{r}
str_view(stringr::words, "ou|ise$|ae|oe|yse$", match = TRUE)
```

## 5 Create a regular expression that will match telephone numbers as commonly written in your country.
```{r}
x <- c("123-456-7890", "(123)456-7890", "(123) 456-7890")
str_view(x, "\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d")
```
```{r}
str_view(x, "\\(\\d\\d\\d\\)\\s*\\d\\d\\d-\\d\\d\\d\\d")
```
## 14.3.4 Repetition
### The next step up in power involves controlling how many times a pattern matches:
### ?: 0 or 1
### +: 1 or more
### *: 0 or more

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
```

```{r}
str_view(x, "CC+")
```
```{r}
str_view(x, 'C[LX]+')
```

### Note that the precedence of these operators is high, so you can write: colou?r to match either American or British spellings. That means most uses will need parentheses, like bana(na)+.

### You can also specify the number of matches precisely:

### {n}: exactly n
```{r}
str_view(x, "C{2}")
```

###{n,}: n or more
```{r}
str_view(x, "C{2,}")
```

### {,m}: at most m
### {n,m}: between n and m


### By default these matches are “greedy”: they will match the longest string possible. You can make them “lazy”, matching the shortest string possible by putting a ? after them. This is an advanced feature of regular expressions, but it’s useful to know that it exists:
```{r}
str_view(x, 'C{2,3}?')
```
```{r}
str_view(x, 'C[LX]+?')
```

## 14.3.4.1 Exercises
## 1. Describe the equivalents of ?, +, * in {m,n} form.
### ? is equivalent to {0,1}, + is equivalent to {1,}, and * is equivalent to {0,}
## 2. Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)
### ^.*$ is a regular expression that will match any string
### "\\{.+\\}" is a string that defines a regular expression that will match any string with {} around 1 or more characters. 
### \d{4}-\d{2}-\d{2} is a regular expression that will match any four digits - 2 digits - 2 digits which is the typical format for a date. 
### "\\\\{4}" is a string that defines a regular expression that matches 4 backlashes. 

## 3.Create regular expressions to find all words that:
### Start with three consonants.
```{r}
str_view(words, "^[^aeiou]{3}", match = TRUE)
```
### Have three or more vowels in a row.
```{r}
str_view(words, "[aeiou]{3,}", match = TRUE)
```

### Have two or more vowel-consonant pairs in a row.
```{r}
str_view(words, "([aeiou][^aeiou]){2,}", match = TRUE)
```

## 14.3.5 Grouping and backreferences
### A capturing group stores the part of the string matched by the part of the regular expression inside the parentheses. You can refer to the same text as previously matched by a capturing group with backreferences, like \1, \2 etc. For example, the following regular expression finds all fruits that have a repeated pair of letters.
```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

## 14.3.5.1 Exercises
## 1.  Describe, in words, what these expressions will match:
### (.)\1\1 will match any word with 3 repeating characters.
### "(.)(.)\\2\\1" will match a pair of characters followed by the same pair in reverse like "abba".
### (..)\1 will match any two characters repeated once
### "(.).\\1.\\1" (original character)(any character)(original character)(any character)(original character)
### "(.)(.)(.).*\\3\\2\\1" (3 characters)(0 or more of any character(s))(original 3 characters in reverse order)
## 2. Construct regular expressions to match words that:
### Start and end with the same character.
```{r}
str_view(words, "^(.)((.*\\1$)|\\1?$)", match = TRUE) 
```
### Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
```{r}
str_view(words, "([A-Za-z][A-Za-z]).*\\1", match = TRUE)
```

## Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)
```{r}
str_view(words, "([a-z]).*\\1.*\\1", match = TRUE)
```
# 14.4 Tools

## 14.4.1 Detect matches
### To determine if a character vector matches a pattern, use str_detect(). It returns a logical vector the same length as the input:

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```
### Remember that when you use a logical vector in a numeric context, FALSE becomes 0 and TRUE becomes 1. That makes sum() and mean() useful if you want to answer questions about matches across a larger vector:
### How many common words start with t?
```{r}
sum(str_detect(words, "^t"))
```
### What proportion of common words end with a vowel?
```{r}
mean(str_detect(words, "[aeiou]$"))
```

### When you have complex logical conditions (e.g. match a or b but not c unless d) it’s often easier to combine multiple str_detect() calls with logical operators, rather than trying to create a single regular expression. For example, here are two ways to find all words that don’t contain any vowels:
### Find all words containing at least one vowel, and negate
```{r}
no_vowels_1 <- !str_detect(words, "[aeiou]")
```

### Find all words consisting only of consonants (non-vowels)
```{r}
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)
```

### A common use of str_detect() is to select the elements that match a pattern. You can do this with logical subsetting, or the convenient str_subset() wrapper:
```{r}
words[str_detect(words, "x$")]
```
```{r}
str_subset(words, "x$")
```

### Typically, however, your strings will be one column of a data frame, and you’ll want to use filter instead:
```{r}
df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(word, "x$"))
```
### A variation on str_detect() is str_count(): rather than a simple yes or no, it tells you how many matches there are in a string:
```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
```
### # On average, how many vowels per word?
```{r}
mean(str_count(words, "[aeiou]"))
```

### It’s natural to use str_count() with mutate():
```{r}
df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )
```

### Note that matches never overlap. For example, in "abababa", how many times will the pattern "aba" match? Regular expressions say two, not three:
```{r}
str_count("abababa", "aba")
```
```{r}
str_view_all("abababa", "aba")
```

## 14.4.1.1 Exercises
## 1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.
### Find all words that start or end with x.
```{r}
words[str_detect(words, "^x|x$")]
```

### Find all words that start with a vowel and end with a consonant.
```{r}
str_subset(words, "^[aeiou].*[^aeiou]$") %>% head()
```

### Are there any words that contain at least one of each different vowel?
```{r}
words[str_detect(words, "a") &
  str_detect(words, "e") &
  str_detect(words, "i") &
  str_detect(words, "o") &
  str_detect(words, "u")]
```

## 2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)
### The word with the highest number of vowels is
```{r}
vowels <- str_count(words, "[aeiou]")
words[which(vowels == max(vowels))]
```

```{r}
prop_vowels <- str_count(words, "[aeiou]") / str_length(words)
words[which(prop_vowels == max(prop_vowels))]
```

# 14.4.2 Extract matches
### To extract the actual text of a match, use str_extract(). 
```{r}
length(sentences)
```
```{r}
head(sentences)
```


### Imagine we want to find all sentences that contain a colour. We first create a vector of colour names, and then turn it into a single regular expression:
```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

### Now we can select the sentences that contain a colour, and then extract the colour to figure out which one it is:
```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

```{r}
has_colour
```

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)
```
```{r}
str_extract(more, colour_match)
```

### To get all matches, use str_extract_all(). It returns a list:

```{r}
str_extract_all(more, colour_match)
```

### If you use simplify = TRUE, str_extract_all() will return a matrix with short matches expanded to the same length as the longest:

```{r}
str_extract_all(more, colour_match, simplify = TRUE)
```

```{r}
x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

## 14.4.2.1 Exercises
## 1. 
```{r}
colour_match2 <- str_c("\\b(", str_c(colours, collapse = "|"), ")\\b")
colour_match2
```
```{r}
more2 <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more2, colour_match2, match = TRUE)
```
## 2. From the Harvard sentences data, extract:
### The first word from each sentence.
```{r}
str_extract(sentences, "[A-Za-z][A-Za-z']*") %>% head()
```
### All words ending in ing.
```{r}
pattern <- "\\b[A-Za-z]+ing\\b"
sentences_with_ing <- str_detect(sentences, pattern)
unique(unlist(str_extract_all(sentences[sentences_with_ing], pattern))) %>%
  head()
```

### All plurals.
```{r}
unique(unlist(str_extract_all(sentences, "\\b[A-Za-z]{3,}s\\b"))) %>%
  head()
```

# 14.4.3 Grouped matches
### You can also use parentheses to extract parts of a complex match. For example, imagine we want to extract nouns from the sentences. As a heuristic, we’ll look for any word that comes after “a” or “the”. 
```{r}
noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)
```

### str_extract() gives us the complete match; str_match() gives each individual component. Instead of a character vector, it returns a matrix, with one column for the complete match followed by one column for each group:
```{r}
has_noun %>% 
  str_match(noun)
```

### If your data is in a tibble, it’s often easier to use tidyr::extract(). It works like str_match() but requires you to name the matches, which are then placed in new columns:
```{r}
tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )
```

## 14.4.3.1 Exercises
## 1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
```{r}
numword <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten) +(\\w+)"
sentences[str_detect(sentences, numword)] %>%
  str_extract(numword)
```
##2. Find all contractions. Separate out the pieces before and after the apostrophe.
```{r}
contraction <- "([A-Za-z]+)'([A-Za-z]+)"
```
```{r}
sentences[str_detect(sentences, contraction)] %>%
  str_extract(contraction) %>%
  str_split("'")
```

## 14.4.4 Replacing matches
### str_replace() and str_replace_all() allow you to replace matches with new strings. The simplest use is to replace a pattern with a fixed string:
```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
```
```{r}
str_replace_all(x, "[aeiou]", "-")
```
```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```
### Instead of replacing with a fixed string you can use backreferences to insert components of the match. In the following code, the order of the second and third words is flipped.
```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

## 14.4.4.1 Exercises
## Replace all forward slashes in a string with backslashes.
```{r}
str_replace_all("past/present/future", "/", "\\\\")
#> [1] "past\\present\\future"
```
## Implement a simple version of str_to_lower() using replace_all().
```{r}

```
```{r}

```

## Switch the first and last letters in words. Which of those strings are still words?
```{r}
swapped <- str_replace_all(words, "^([[:alpha:]])(.*)([[:alpha:]])$", "\\3\\2\\1")
```
```{r}
intersect(swapped, words)
```

# 14.4.5 Splitting
### split sentences into words:
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```
```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```

## Otherwise, like the other stringr functions that return a list, you can use simplify = TRUE to return a matrix:
```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

## You can also request a maximum number of pieces:
```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

## Instead of splitting up strings by patterns, you can also split up by character, line, sentence and word boundary()s:
```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

```{r}
str_split(x, " ")[[1]]
```

```{r}
str_split(x, boundary("word"))[[1]]
```

## 14.4.5.1 Exercises

## 1. Split up a string like “apples, pears, and bananas” into individual components.
```{r}
x <- c("apples, pears, and bananas")
```
```{r}
str_split(x, ", +(and +)?")[[1]]
```


## 2. Why is it better to split up by boundary(“word”) than " "?
### Splitting by boundary("word") is better than "" because it recognizes punctation that splits words and does not add extra spaces. 

## 3. What does splitting with an empty string ("") do? Experiment, and then read the documentation.
```{r}
str_split("ab. cd|agt", "")[[1]]
#>  [1] "a" "b" "." " " "c" "d" "|" "a" "g" "t"
```
















