---
title: "Lab_6_Covid_Report"
author: "Ben Earle"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
install.packages("wesanderson")
```

```{r, include=FALSE}
library(wesanderson)
```
```{r, include=FALSE}
install.packages("maps")
```

```{r, include=FALSE}
install.packages("mapdata")
```

```{r, include=FALSE}
library(maps)
```

```{r, include=FALSE}
library(mapdata)
```

```{r, include=FALSE}
install.packages("viridis")
```

```{r, include=FALSE}
library(viridisLite)
```


```{r, include=FALSE}
library(viridis)
```

```{r, include=FALSE}
library(lubridate)
```
```{r, include=FALSE}
library(tidyverse)
```


### Covid-19 Static Maps and Interactive graphs - April 2nd vs. Spetember 26 2020

## The following Map diplays the relative number of confirmed cases on April 2nd

```{r, include=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") 
```

```{r, echo=FALSE}
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed/1000)) +
    borders("world", colour = NA, fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
    labs(title = 'World COVID-19 Confirmed cases',x = '', y = '',
        size="Cases (x1000))") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

## Covid-19 Confirmed Case number as of April 5th, 2020
```{r, include=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
```

```{r, echo=FALSE}
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US', x = '', y = '',
        size="Cases") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

## The following graph gives us an idea where the number of confirmed cases initially were highly concentrated in the US on April 2nd, 2020.

```{r, include=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))
```

```{r, echo=FALSE}
us <- map_data("state")
state_join <- left_join(us, daily_report, by = c("region" = "Province_State"))
ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in the US'")
```

## Here is another look at the number of confirmed cases in US as of April 2nd.
```{r, include=FALSE}
library(RColorBrewer)
```

```{r, include=FALSE}
report_03_27_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  unite(Key, Admin2, Province_State, sep = ".") %>% 
  group_by(Key) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Key = tolower(Key))
```

```{r, echo=FALSE}
us <- map_data("state")
counties <- map_data("county") %>% 
  unite(Key, subregion, region, sep = ".", remove = FALSE)
state_join <- left_join(counties, report_03_27_2020, by = c("Key"))

ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  borders("state", colour = "black") +
  geom_polygon(data = state_join, aes(fill = Confirmed)) +
  scale_fill_gradientn(colors = brewer.pal(n = 5, name = "PuRd"),
                       breaks = c(1, 10, 100, 1000, 10000, 100000),
                       trans = "log10", na.value = "White") +
  ggtitle("Number of Confirmed Cases by US County") +
  theme_bw() 
```

## Here is a look at where the number of confirmed cases of Covid were intially the highest on April 2nd.

```{r, include=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-02-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
state_join <- left_join(ma_county, daily_report, by = c("subregion" = "Admin2")) 
```

```{r, echo=FALSE}
ggplot(data = ma_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) +
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "white") +
    scale_fill_gradientn(colors = brewer.pal(n = 5, name = "BuGn"),
                         trans = "log10") +
  labs(title = "COVID-19 Confirmed Cases in Massachusetts'")
```

## Now, let's take a look at the number of deaths throughout the entire world as of septmember 26th

```{r, include=FALSE}
# Read in the daily report
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  group_by(Country_Region) %>% 
  summarize(Confirmed = sum(Confirmed), Deaths = sum(Deaths))

# Read in the world map data
world <- as_tibble(map_data("world"))

# Check to see if there are differences in the naming of countries
setdiff(world$region, daily_report$Country_Region)
```

```{r, echo=FALSE}
world <- as_tibble(map_data("world")) %>% 
 mutate(region = str_replace_all(region, c("USA" = "US", "Czech Republic" = "Czechia",  
        "Ivory Coast" = "Cote d'Ivoire", "Democratic Republic of the Congo" = "Congo (Kinshasa)", 
        "Republic of Congo" = "Congo (Brazzaville)"))) 

# Join the covid report with the map data
country_join <- left_join(world, daily_report, by = c("region" = "Country_Region"))

# Create the graph
library(plotly)
ggplotly(
ggplot(data = world, mapping = aes(x = long, y = lat, text = region, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = country_join, aes(fill = Deaths), color = "black") +
  scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  labs(title = "COVID-19 Deaths")
) 
```

## Locations of confirmed cases in the US as of Septemebr 26th, 2020. As you can see, there is a drastically higher number of cases of the right side of the country where population density is higher.

```{r, include=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  filter (!Province_State %in% c("Alaska","Hawaii", "American Samoa",
                  "Puerto Rico","Northern Mariana Islands", 
                  "Virgin Islands", "Recovered", "Guam", "Grand Princess",
                  "District of Columbia", "Diamond Princess")) %>% 
  filter(Lat > 0)
```

```{r, echo=FALSE}
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "black", fill = "grey90") +
    theme_bw() +
    geom_point(shape = 21, color='purple', fill='purple', alpha = 0.5) +
    labs(title = 'COVID-19 Confirmed Cases in the US', x = '', y = '',
        size="Cases") +
    theme(legend.position = "right") +
    coord_fixed(ratio=1.5)
```

## Here is another look at the same graph with relative density of cases represented with color.

```{r}
mybreaks <- c(1, 100, 1000, 10000, 10000)
ggplot(daily_report, aes(x = Long, y = Lat, size = Confirmed)) +
    borders("state", colour = "white", fill = "grey90") +
    geom_point(aes(x=Long, y=Lat, size=Confirmed, color=Confirmed),stroke=F, alpha=0.7) +
    scale_size_continuous(name="Cases", trans="log", range=c(1,7), 
                        breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+")) +
    scale_color_viridis_c(option="viridis",name="Cases", trans= "log", breaks=mybreaks, labels = c("1-99",
                        "100-999", "1,000-9,999", "10,000-99,999", "50,000+"))  +
# Cleaning up the graph
  
  theme_void() + 
    guides( colour = guide_legend()) +
    labs(title = "Anisa Dhana's lagout for COVID-19 Confirmed Cases in the US'") +
    theme(
      legend.position = "bottom",
      text = element_text(color = "#22211d"),
      plot.background = element_rect(fill = "#ffffff", color = NA), 
      panel.background = element_rect(fill = "#ffffff", color = NA), 
      legend.background = element_rect(fill = "#ffffff", color = NA)
    ) +
    coord_fixed(ratio=1.5)
```

## Here is a map displaying the relative density of confirmed cases in California as of September 26th, 2020

```{r, include=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-26-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "California") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))
us <- map_data("state")
ca_us <- subset(us, region == "california")
counties <- map_data("county")
ca_county <- subset(counties, region == "california")
state_join <- left_join(ca_county, daily_report, by = c("subregion" = "Admin2")) 
```

```{r, echo=FALSE}
library(wesanderson)
ggplotly(
  ggplot(data = ca_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
# Add data layer
  geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                         wes_palette("Zissou1", 100, type = "continuous")) +
  ggtitle("COVID-19 Cases in CA") +
# Cleaning up the graph
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank()) +
  theme(panel.background = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_blank())
)
```

# REFERENCES

## 1
title: "Evolutionary Genomics & Bioinformatics"
author: "Jeffrey Blanchard"
date: "2/26/2020"
output: html_document
bibliography: data/references.bib

Application written in R [@RCoreTeam] using the Shiny framework [@Chang2015].

## 2
title: "Draw Geographical Maps"
author: "Eric Anderson"
date: "4/03/2018"
output: pdf_document
bibliography: https://cran.r-project.org/web/packages/maps/maps.pdf

##3 
title: "An interactive web-based dashboard to track COVID-19 in real time"
author: "Ensheng Dong, Hongru Du, Lauren Gardner"
date: "2/19/2020"
output: html_document
bibliography: https://www.sciencedirect.com/science/article/pii/S1473309920301201?via%3Dihub

